Sending a Balloon to the Edge of Space

### 2018-11-14 paper/final_comp.Rmd
edits
more edits

---
\includepdf[pages={-}]{paper.pdf}
\includepdf[pages={-}]{bibliography.pdf}
\includepdf[pages={-}]{bibliography.pdf}

@@

```{r appendix2, echo=F}
knitr::include_graphics("assets/appendix2.png")
```

@@

# Appendix III: Layers of the Atmosphere
```{r appendix3, echo=F}
knitr::include_graphics("assets/appendix3.png")
```

\includepdf[pages={-}]{assets/cad1.pdf}
\includepdf[pages={-}]{assets/cad2.pdf}
\includepdf[pages={-}]{assets/cad3.pdf}
\includepdf[pages={-}]{assets/cad4.pdf}

>>>>

---
\includepdf[pages={-}]{paper.pdf}
\includepdf[pages={-}]{bibliography.pdf}
\includepdf[pages={-}]{assets/ap1.pdf}

@@

```{r appendix2, echo=F, out.width="400px"}
knitr::include_graphics("assets/appendix2.png")
```

@@

# Appendix III: Layers of the Atmosphere
```{r appendix3, echo=F, out.width="400px"}
knitr::include_graphics("assets/appendix3.png")
```

\includepdf[pages={-}]{assets/cad1.pdf}
\includepdf[pages={-}]{assets/cad2.pdf}
\includepdf[pages={-}]{assets/cad3.pdf}
\includepdf[pages={-}]{assets/cad4.pdf}
\includepdf[pages={-}]{assets/Yagi.pdf}
\includepdf[pages={-}]{assets/reel_in.pdf}
\includepdf[pages={-}]{assets/schematic.pdf}

# Appendix IX: Modeled vs. Actual Trajectory

```{r kml, echo=F, out.width="400px"}
knitr::include_graphics("assets/kml.png")
```

Here is an overlay of the simulated balloon trajectory (taken 12 hours before the launch) and the actual recorded balloon trajectory. The trajectory in the back is the actual data, and the trajectory in front is the simulated. There is a gap in the actual trajectory, which is due to the fact that the logger stopped recording data midway through its flight. In addition, because the balloon popped earlier than anticipated, it was not able to gain the expected altitude of 33km so it landed closer to the launch site. However, had conditions been optimal and the balloon reached its optimal altitude, the simulation probably would have been a good predictor of actual balloon trajectory.

### 2018-11-14 paper/paper.rmd
more edits

@@

@@

The cutdown was tested on a seperate arduino in a vacuum chamber. The code used for this test follows the averaging method described above and can be found in the appendix. The cutdown successfully burnt through a string to release a 1.4 kg weight when the vacuum chamber dropped below 1.08 kPa. 1.08 kPa is the cutdown pressure specified in the code. This value was chosen because it was the pressure at 100,000 ft.

@@

As discussed, the launch can be considered a technical success. With the payload and most data recovered, atmospheric analysis has been made possible by 6thsense's launch. The results discussed above agree with the scientific community's current understanding of the atmosphere and atmospheric layers. Not only was data recorded to study the atmosphere, however, but 6thsense's balloon launch was also a test of systems that the team designed to meet the technical challenges of sending equipment to the edge of space.

The 6thsense team, working together for the first time on such a large project, grew not only their knowledge of the subject matter required to develop and launch such a payload, but also communication and collaboration skills, necessary in any large scall venture.

>>>>

In addition to information about temperature, NASA’s empirical model also accounts for the relationship between air pressure and time. Since atmospheric composition and temperature vary, a simple exponential model is not enough to correctly predict the relationship between pressure and height. For example, the temperature stops changes in most parts of the atmosphere and remains constant in the lower stratosphere. A simple exponential model would not account for this variation in temperature change. NASA’s empirical model separates fitted curves to account for differences in atmospheric compositions in order to derive a more accurate relationship between pressure and altitude.

@@

## Calculating Payload Weight

### Buoyancy Force Calculations

$$F_B=\rho g V$$
$$\rho = 1.225kg/m^3, \, g = 9.8m/s^2, \, V = \frac{4}{3}\pi(0.8382m)^3=2.87m^3$$
$$F_B = (1.225kg/m^3)(9.8m/s^2)(2.87m^3)= 37.2N$$
$$F_{net} = F_B - F_{weight} = (mass)(acceleration)$$
$$let\, \,\,mass = m_{system}, \, F_B = 37.2N, \, F_{weight} = mg = m_{system}(9.8m/s^2)$$
$$a = \frac{37.2 - 9.8m_{system}}{m_{system}}$$
We want to reach an altitude of 33,500 m and come back within 3 hours:
$$v_{avg} = \frac{33500 * 2}{10,800} = 6.2m/s$$
$$\int a\,dt = v_{avg} = 6.2m/s$$
$$\int a\,dt = \int \frac{37.2-9.8m_{system}}{m_{system}} \, dt$$
$$6.2 = \frac{37.2-9.8m_{system}}{m_{system}}$$
$$6.2m_{system} = 37.2 - 9.8m_{payload}$$
$$16m_{system} = 37.2$$
$$\boxed{m_{system} = 2.325kg}$$
$$m_{system} = m_{balloon} + m_{payload}, \, \, m_{balloon} \approx 1kg$$
$$m_{payload} \approx 1.3kg$$

### Buoyancy Force at Highest Altitude

$$F_B = \rho_{air} V_{balloon} g$$
$$PV=nRT$$
$$V_{balloon}=\frac{nRT}{P}$$

$$n=128.125mol, \, R = 0.8206, \, T=223K, \, P=0.00967atm$$
$$V_{balloon} = \frac{128.125 * 0.8206 *223}{0.00967} = 242462\,\,liters$$
$$V_{balloon} = 242.5m^3$$
$$\rho_{35000m}=0.0082kg/m^3$$
$$\boxed{F_B = (0.0082kg/m^3)(242.5m^3)(9.8m/s^2)=19.49N}$$

### Derivation of Exponential Model of Pressure

$$pressure = \frac{Force}{Area} = \rho gh$$
$$dP=-\rho g\,dh$$
$$PV=nRT$$
$$\frac{n}{v} = \frac{P}{rt}$$

$$\rho = density = \frac{mass}{volume} $$
$$\rho = \frac{nm_{mol}}{v} = \frac{Pm_{mol}}{rt}$$
$$\rho = \frac{P}{Rt}(m_{mol})$$
$$dP = -\frac{Pm_{mol}}{Rt}g\,dh$$
$$\frac{dP}{P} =  -\frac{m_{mol}}{Rt}g\,dh$$
$$\int_{0}^{h} \frac{dP}{P} =- \int_{0}^{h} \frac{m_{mol}}{Rt}g\,dh$$
$$\ln{P} = -\frac{m_{mol}g}{Rt}\,dh$$
$$m_{mol}=0.29\,g/mol, \,g=\,9.8m/s^2, \,R=8.314\,J/mol, \,t=273 \, K$$
$$\frac{m_{mol}g}{Rt} = \frac{1}{8.4km}$$
$$\ln{P} = -\frac{h}{8.4}$$
$$e^{\ln{P}}=P_0 \,e^{\frac{-h}{8.4}}$$
$$\boxed{P = P_0e^{\frac{-h}{8.4}}}$$
$$\frac{P}{P_0} = e^{\frac{-h}{8.4}}$$
$$\ln{(\frac{P}{P_0})} = \ln{e^{\frac{-h}{8.4}}}$$
$$\frac{-h}{8.4} = \ln{(\frac{P}{P_0})}$$
$$h = -8.4 \, \ln{(\frac{P}{P_0})}$$
$$\boxed{h = -a \, \ln{(\frac{P}{P_0})}}$$
The exponential model of pressure only holds up to the edge of the stratosphere, and does not apply to higher altitudes. This is because this model depends on temperature being relatively constant, which is true up to the stratosphere. However, beyond the stratosphere, temperatures are wildyly variable,  and therefore this model no longer holds true.

@@

The cutdown was tested on a seperate arduino in a vacuum chamber. The code used for this test follows the averaging method described above. The cutdown successfully burnt through a string to release a 1.4 kg weight when the vacuum chamber dropped below 1.08 kPa. 1.08 kPa is the cutdown pressure specified in the code. This value was chosen because it was the pressure at 100,000 ft.

@@

As discussed, the launch can be considered a technical success. With the payload and most data recovered, atmospheric analysis has been made possible by 6thsense's launch. The results discussed above agree with the scientific community's current understanding of the atmosphere and atmospheric layers. Not only was data recorded to study the atmosphere, but 6thsense's balloon launch was also a test of systems that the team designed to meet the technical challenges of sending equipment to the edge of space.

The 6thsense team, working together for the first time on such a large project, grew not only their knowledge of the subject matter required to develop and launch such a payload, but also communication and collaboration skills, necessary in any large scale venture.

# Acknowledgements

This project would have been possible without the continued support of Dr. James Dann and Mr. Ward. Their willingness to go out of their way to provide guidance has been extremely helpful greatly appreciated by the entire team.

In addition, we would like to thank all the members of other ASR teams who we collaborated with, whose knowledge we benefited from.

### 2018-11-14 paper/paper.tex
more edits

@@

@@

\begin{center}\includegraphics[width=400px]{assets/schematic} \end{center}

The circuit consists of two Arduinos, one Mega ATK and one Uno. The Mega
ATK is our primary microcontroller and controls the logger, transmitter,
and receives input from every sensor except for the accelerometer. The
accelerometer and another microSD logger are the only two components on
the secondary Arduino Uno. The reason for the accelerometer needing it's
own Arduino is because the \(I^2C\) library it required was not
compatible with the other devices running on \(I^2C\), such as the
BME280 and spectrometer. The primary Arduino is powered by three
batteries, and one battery powers the secondary arduino and the methane
sensor. The methane sensor uses a large amount of power so it is with
the secondary Arduino's power supply so that if the sensor uses too much
power it will not disrupt the majority of the data collection and
sensors. All batteries go through 1 Amp 5 Volt voltage regulators, one
for the primary Arduino and one for the secondary Arduino and methane
sensor. The built in voltage regulator on the Arduino was not powerful
enough and causing the sensors to malfunction so exterior regulators
were included. Every sensor connects individually to either an analog
pin or digital pin, or is one of the three sensors that connect to the
SDA and SCL pins.

@@

As discussed, the launch can be considered a technical success. With the
payload and most data recovered, atmospheric analysis has been made
possible by 6thsense's launch. The results discussed above agree with
the scientific community's current understanding of the atmosphere and
atmospheric layers. Not only was data recorded to study the atmosphere,
however, but 6thsense's balloon launch was also a test of systems that
the team designed to meet the technical challenges of sending equipment
to the edge of space.

The 6thsense team, working together for the first time on such a large
project, grew not only their knowledge of the subject matter required to
develop and launch such a payload, but also communication and
collaboration skills, necessary in any large scall venture.

>>>>

In addition to information about temperature, NASA's empirical model
also accounts for the relationship between air pressure and time. Since
atmospheric composition and temperature vary, a simple exponential model
is not enough to correctly predict the relationship between pressure and
height. For example, the temperature stops changes in most parts of the
atmosphere and remains constant in the lower stratosphere. A simple
exponential model would not account for this variation in temperature
change. NASA's empirical model separates fitted curves to account for
differences in atmospheric compositions in order to derive a more
accurate relationship between pressure and altitude.

@@

\subsection{Calculating Payload
Weight}\label{calculating-payload-weight}

\subsubsection{Buoyancy Force
Calculations}\label{buoyancy-force-calculations}

\[F_B=\rho g V\]
\[\rho = 1.225kg/m^3, \, g = 9.8m/s^2, \, V = \frac{4}{3}\pi(0.8382m)^3=2.87m^3\]
\[F_B = (1.225kg/m^3)(9.8m/s^2)(2.87m^3)= 37.2N\]
\[F_{net} = F_B - F_{weight} = (mass)(acceleration)\]
\[let\, \,\,mass = m_{system}, \, F_B = 37.2N, \, F_{weight} = mg = m_{system}(9.8m/s^2)\]
\[a = \frac{37.2 - 9.8m_{system}}{m_{system}}\] We want to reach an
altitude of 33,500 m and come back within 3 hours:
\[v_{avg} = \frac{33500 * 2}{10,800} = 6.2m/s\]
\[\int a\,dt = v_{avg} = 6.2m/s\]
\[\int a\,dt = \int \frac{37.2-9.8m_{system}}{m_{system}} \, dt\]
\[6.2 = \frac{37.2-9.8m_{system}}{m_{system}}\]
\[6.2m_{system} = 37.2 - 9.8m_{payload}\] \[16m_{system} = 37.2\]
\[\boxed{m_{system} = 2.325kg}\]
\[m_{system} = m_{balloon} + m_{payload}, \, \, m_{balloon} \approx 1kg\]
\[m_{payload} \approx 1.3kg\]

\subsubsection{Buoyancy Force at Highest
Altitude}\label{buoyancy-force-at-highest-altitude}

\[F_B = \rho_{air} V_{balloon} g\] \[PV=nRT\]
\[V_{balloon}=\frac{nRT}{P}\]

\[n=128.125mol, \, R = 0.8206, \, T=223K, \, P=0.00967atm\]
\[V_{balloon} = \frac{128.125 * 0.8206 *223}{0.00967} = 242462\,\,liters\]
\[V_{balloon} = 242.5m^3\] \[\rho_{35000m}=0.0082kg/m^3\]
\[\boxed{F_B = (0.0082kg/m^3)(242.5m^3)(9.8m/s^2)=19.49N}\]

\subsubsection{Derivation of Exponential Model of
Pressure}\label{derivation-of-exponential-model-of-pressure}

\[pressure = \frac{Force}{Area} = \rho gh\] \[dP=-\rho g\,dh\]
\[PV=nRT\] \[\frac{n}{v} = \frac{P}{rt}\]

\[\rho = density = \frac{mass}{volume} \]
\[\rho = \frac{nm_{mol}}{v} = \frac{Pm_{mol}}{rt}\]
\[\rho = \frac{P}{Rt}(m_{mol})\] \[dP = -\frac{Pm_{mol}}{Rt}g\,dh\]
\[\frac{dP}{P} =  -\frac{m_{mol}}{Rt}g\,dh\]
\[\int_{0}^{h} \frac{dP}{P} =- \int_{0}^{h} \frac{m_{mol}}{Rt}g\,dh\]
\[\ln{P} = -\frac{m_{mol}g}{Rt}\,dh\]
\[m_{mol}=0.29\,g/mol, \,g=\,9.8m/s^2, \,R=8.314\,J/mol, \,t=273 \, K\]
\[\frac{m_{mol}g}{Rt} = \frac{1}{8.4km}\] \[\ln{P} = -\frac{h}{8.4}\]
\[e^{\ln{P}}=P_0 \,e^{\frac{-h}{8.4}}\]
\[\boxed{P = P_0e^{\frac{-h}{8.4}}}\]
\[\frac{P}{P_0} = e^{\frac{-h}{8.4}}\]
\[\ln{(\frac{P}{P_0})} = \ln{e^{\frac{-h}{8.4}}}\]
\[\frac{-h}{8.4} = \ln{(\frac{P}{P_0})}\]
\[h = -8.4 \, \ln{(\frac{P}{P_0})}\]
\[\boxed{h = -a \, \ln{(\frac{P}{P_0})}}\] The exponential model of
pressure only holds up to the edge of the stratosphere, and does not
apply to higher altitudes. This is because this model depends on
temperature being relatively constant, which is true up to the
stratosphere. However, beyond the stratosphere, temperatures are wildyly
variable, and therefore this model no longer holds true.

@@

The circuit consists of two Arduinos, one Mega ATK and one Uno. The Mega
ATK is our primary microcontroller and controls the logger, transmitter,
and receives input from every sensor except for the accelerometer. The
accelerometer and another microSD logger are the only two components on
the secondary Arduino Uno. The reason for the accelerometer needing it's
own Arduino is because the \(I^2C\) library it required was not
compatible with the other devices running on \(I^2C\), such as the
BME280 and spectrometer. The primary Arduino is powered by three
batteries, and one battery powers the secondary arduino and the methane
sensor. The methane sensor uses a large amount of power so it is with
the secondary Arduino's power supply so that if the sensor uses too much
power it will not disrupt the majority of the data collection and
sensors. All batteries go through 1 Amp 5 Volt voltage regulators, one
for the primary Arduino and one for the secondary Arduino and methane
sensor. The built in voltage regulator on the Arduino was not powerful
enough and causing the sensors to malfunction so exterior regulators
were included. Every sensor connects individually to either an analog
pin or digital pin, or is one of the three sensors that connect to the
SDA and SCL pins. See Appendix for schematic.

@@

As discussed, the launch can be considered a technical success. With the
payload and most data recovered, atmospheric analysis has been made
possible by 6thsense's launch. The results discussed above agree with
the scientific community's current understanding of the atmosphere and
atmospheric layers. Not only was data recorded to study the atmosphere,
but 6thsense's balloon launch was also a test of systems that the team
designed to meet the technical challenges of sending equipment to the
edge of space.

The 6thsense team, working together for the first time on such a large
project, grew not only their knowledge of the subject matter required to
develop and launch such a payload, but also communication and
collaboration skills, necessary in any large scale venture.

\section{Acknowledgements}\label{acknowledgements}

This project would have been possible without the continued support of
Dr.~James Dann and Mr.~Ward. Their willingness to go out of their way to
provide guidance has been extremely helpful greatly appreciated by the
entire team.

In addition, we would like to thank all the members of other ASR teams
who we collaborated with, whose knowledge we benefited from.
